A Survey of Physical Design Techniques of Information Systems

### 2010-07-26 cs848-db/Paper.tex
after integrating everything

\subsubsection*{B Trees}
Although B-trees, explained in the previous section, are originally designed for disk based database systems, they were also found to be useful in
MMDB~\cite{lehman1986study}. In disk based systems, the B+ Tree is more used than the regular B Tree. However, in MMDBs, the B-tree would be
more suitable from a storage perspective since there is no gain from keeping all the data in the leaves. In a main memory system, this
would only waste space without improving performance. On the other hand, the B+ Tree uses multiple keys to search within a node. This means that
if the node fits in a cache line, this cache load can satisfy more than more comparison leading to better cache utilization~\cite{rao1999cache}. Additionally,
less data per node is stored in a B+tree which means there is better cache utilization. Therefore, B+trees in general have reasonable cache behavior, and
variations of them have been adapted in MMDBs.

@@

\subsubsection{Materialized Views}
not very useful since most queries are ad-hoc \cite{armstrong1997data, theodoratos2000general}

\subsubsection{Partitioning}
\cite{thusoo2010data}
\paragraph{Vertical Partitioning}

\paragraph{Horizontal Partitioning}

@@

Automating the physical database design choices has been a very attractive area of research that is still receiving lots of attention until now. Choosing which
indexes to create, which views to materialize, and how to partition data are hard choices. They are highly dependent on the schema used as well
as the query workload. The number of combinations of choices that can be made is very large making this task quite tedious and difficult for database
administrators. Accordingly, researchers have been inclined to find ways to automate these decisions. There were many early research efforts done in this
area(E.g.~\cite{finkelstein1988physical, hammer76selec, frank1992selec, chaudhuri1997overview, schiefer1999db2, valentin2000select}), but most of these efforts
focused on having a static workload. Accordingly, research shifted to more dynamic ways to automate physical design. Initial attempts used an dynamic, on-line
approaches(E.g.~\cite{bruno2007online, sattler2003, schnaitter2007}) where they observed the changes in the workload and adjusted the design accordingly while
attempting to predict future behavior based on the observed behavior. To avoid the problems of on-line design, researchers are also currently looking at how to
perform this dynamic physical design off-line (E.g. ~\cite{hannes2008,agrawal2006}). Additionally, finding better evaluation techniques to compare different
automatic physical design decision is currently being studied~\cite{gebaly2008}.

Other problems which are currently being researched include improving I/O operations in database systems by using Solid State Drives (SSD) instead of of Hard
Disk Drives (HDD)~\cite{lee2009, du2009, lee2008}. SSDs have faster read performance, but perform poorly in random writes. However, for certain applications in
databases and for databases where random writes are not as frequent, SSDs might be suitable. Using SDDs also seems promising when used for
transaction logs, and intermediate results of sorting and joining~\cite{lee2008}.

Just like all other research areas are moving to the cloud, database systems are moving there too. Accordingly, lots of current research focuses on adapting
storage techniques and index structures to the cloud. For example, Wang et al. propose a new index structure for indexing multi-dimensional data in the
cloud~\cite{wang2010indexing}. Lots of work is currently being done in this area. For example, Wu et al.~\cite{wu2010efficient} propose a modified B+tree index
for datasets stored in each compute node. The exact requirements and solutions for data management in the cloud are still being
researched~\cite{abadi2009data}, and so this is an area that is expected to receive more attention in the future as cloud computing becomes more prevalent.

@@

\begin{table}[!t]
\centering
\scalebox{0.7}{
\begin{tabular}{p{2cm}||p{5cm}|p{4cm}|p{4cm}|p{4cm}}
 & \textbf{Index Structures} & \textbf{Materialized Views} & \textbf{Partitioning} & \textbf{Clustering}\\
\hline\hline
\textbf{RDBMS}& E.g. B+trees, Bitmap Index, Hash Tree & Store results of common complicated queries on disk & Horizontal Partitioning (range, list, hash,
composite) Vertical Partitioning& Group related items together on disk\\
\hline
\textbf{MMDB} & 
-No need to store actual data values in index \linebreak \linebreak 
-Larger node size that is aligned with cache line size \linebreak \linebreak
-E.g. B+tree, CSB+tree, T tree, CST tree & Not highly utilized since processing and main memory access is cheap & Only needed in secondary storage to speed up
reloading in case of a crash\linebreak \linebreak 
-Horizontal \& Vertical partitioning& 
Not needed since random access in main memory costs the same as sequential access\\
\hline
\textbf{XML DB} & Need a number scheme. E.g. B+tree and variations used. & XQuery and XPath results are materialized & Horizontal partitioning based on inlining
or node
type & Based on document order or structural similarity\\
\hline
\textbf{Data Warehouses} & E.g. Universal B+tree, Bitmap index, Projection Index & 
-Essential to improve performance \linebreak \linebreak 
-Not applicable to adhoc queries & 
-Horizontal partitioning based on time dimension \linebreak \linebreak 
-Vertical partitioning of seldom used columns. & Data is
clustered by
nature.\\
\end{tabular}
}
\caption{Summary of the different elements of physical design in the studied systems}
\label{tab:summary}
\end{table}

@@

In this survey, we present four elements of physical design in database: index structures, materialized views, partitioning, and clustering. We started out
by explaining how these design factors are implemented in traditional Relational Database Management Systems (RDBMS), and then presented how they are used in
other information systems. We focused on three types of information systems: Main Memory Database Systems (MMDB), XML Databases, and Data Warehouses. We
presented the challenges involved with each type of system, and how each of the four physical design elements were adapted or utilized to overcome these
challenges. Table~\ref{tab:summary} summarizes the results of our survey. From our literature review, we found that the B+tree is the most utilized index
structure, and has been adapted and customized to fit the different needs of other systems. For example, it was made more cache conscious to fit MMDB needs,
and was changed to apply to multidimensions to fit the Data Warehouses needs. Materialized views were found to be heavily utlitized in all the systems we
presented, except in MMDBs. Partitioning was also used in all systems, but was only used for data reloading in MMDBs. Horizontal partitioning is much more
common that vertical partitioning. Clustering is very important in all systems, except for MMDBs. Clustering is especially very important for Data Warehouses
where huge amounts of data are stored.

@@

>>>>

\subsubsection*{B-trees}
Although B-trees, explained in the previous section, are originally designed for disk based database systems, they were also found to be useful in
MMDB~\cite{lehman1986study}. In disk based systems, the B+ Tree is more used than the regular B Tree. However, in MMDBs, the B-tree would be
more suitable from a storage perspective since there is no gain from keeping all the data in the leaves. In a main memory system, this
would only waste space without improving performance. On the other hand, the B+ Tree uses multiple keys to search within a node. This means that
if the node fits in a cache line, this cache load can satisfy more than more comparison leading to better cache utilization~\cite{rao1999cache}. Additionally,
less data per node is stored in a B+tree which means there is better cache utilization. Therefore, B+trees in general have reasonable cache behavior, and
variations of them have been adapted in MMDBs.

@@

\subsubsection{Materialized Views}
A data warehouse must be built with a solid foundation that is flexible enough and responsive to business changes over time \cite{armstrong1997data}. Meanwhile,
a data warehouse can be seen as a set of materialized views defined over the remote data sources (Figure \ref{fig:dw}). That is why selection of materialized
views is one of the most challenging problems that faces data warehouse architects \cite{theodoratos2000general}. Most of the solutions in the literature
defined the problem of selecting views to materialize as \textquotedblleft selecting views for materialization in order to satisfy a number of design goals
given a set of constraints \textquotedblright. The problem can be reduced to a cost minimization problem, whether it is the query evaluation cost
\cite{harinarayan1996implementing}, total view maintenance cost \cite{kenneth1996materialized}, both \cite{yang1997algorithms}, or the operational cost
\cite{elena1997materialized, gupta2005selection, theodoratos1999designing, theodoratos1997data, theodoratos1999designingdw, theodoratos2000incremental,
yang1997algorithms}. Although materialized views greatly improves the performance of a data warehouse, it might not be as useful when the query workload is
formed of ad-hoc complex queries.

\subsubsection{Partitioning}
Both horizontal and vertical partitioning are heavily used in data warehouse design. Partitioning a data warehouse has many advantages. Horizontal partitioning
is usually used in data warehouses to make the job of the data warehouse administrators easier by facilitating the process of manages such large data
sources from which a data warehouse gets its data. In other words, it breaks down the management of large databases to managing smaller sets (i.e. partitions)
of data \cite{thusoo2010data}. More importantly, partitioning provides parallelism in various situations: loading data sources \cite{barclay1994loading},
performing parallel joins on huge data \cite{datta2002parallel}, faster drill-down and roll-up operations \cite{chaudhuri1997overview}. Partitioning also
facilitates compressing data that is not of interest to users of the data warehouse. Vertical paritioning is also used to keep seldom used columns away from
the process of answering queries for better performance.

On the other hand, if partitioning is over-used in a data warehouse, managing them will be a complex process due to the fact of having a huge amount of data.
Finally, there is an inherent cost of managing the meta data (e.g. data cube definition) when the number of partitions increases.

\subsubsection{Clustering}
A data warehouse is clustered by nature as Figure \ref{fig:dw} suggests. The data loaded into a data warehouse comes from various operational data sources,
each source has the data that is the most interesting/related to it (i.e. clustering). The data warehouse can be further divided into a collection of
conforming data marts \cite{sen2005comparison}. Each data mart stores data that is related to each other (e.g. based on the organization's business processes).
For example, one data mart can store sales data that is of particular interest to the Sales Department in an organization. Along with sales data,
product-related information can also be stored to provide better locality for queries that usually ask question about sales and products, and so on.

@@

Automating the physical database design choices has been a very attractive area of research that is still receiving lots of attention until now. Choosing which
indexes to create, which views to materialize, and how to partition data are hard choices. They are highly dependent on the schema used as well
as the query workload. The number of combinations of choices that can be made is very large making this task quite tedious and difficult for database
administrators. Accordingly, researchers have been inclined to find ways to automate these decisions. There were many early research efforts done in this
area(E.g. \cite{finkelstein1988physical, hammer76selec, frank1992selec, chaudhuri1997overview, schiefer1999db2, valentin2000select}), but most of these efforts
focused on having a static workload. Accordingly, research shifted to more dynamic ways to automate physical design. Initial attempts used an dynamic, on-line
approaches(E.g. \cite{bruno2007online, sattler2003, schnaitter2007}) where they observed the changes in the workload and adjusted the design accordingly while
attempting to predict future behavior based on the observed behavior. To avoid the problems of on-line design, researchers are also currently looking at how to
perform this dynamic physical design off-line (E.g. \cite{hannes2008,agrawal2006}). Additionally, finding better evaluation techniques to compare different
automatic physical design decision is currently being studied \cite{gebaly2008}.

Other problems which are currently being researched include improving I/O operations in database systems by using Solid State Drives (SSD) instead of of Hard
Disk Drives (HDD) \cite{lee2009, du2009, lee2008}. SSDs have faster read performance, but perform poorly in random writes. However, for certain applications in
databases and for databases where random writes are not as frequent, SSDs might be suitable. Using SDDs also seems promising when used for
transaction logs, and intermediate results of sorting and joining \cite{lee2008}.

Just like all other research areas are moving to the cloud, database systems are moving there too. Accordingly, lots of current research focuses on adapting
storage techniques and index structures to the cloud. For example, Wang et al. propose a new index structure for indexing multi-dimensional data in the
cloud \cite{wang2010indexing}. Lots of work is currently being done in this area. For example, Wu et al. \cite{wu2010efficient} propose a modified B+tree index
for datasets stored in each compute node. The exact requirements and solutions for data management in the cloud are still being
researched \cite{abadi2009data}, and so this is an area that is expected to receive more attention in the future as cloud computing becomes more prevalent.

@@

@@

In this survey, we present four elements of physical design in database: index structures, materialized views, partitioning, and clustering. We started out
by explaining how these design factors are implemented in traditional Relational Database Management Systems (RDBMS), and then presented how they are used in
other information systems. We focused on three types of information systems: Main Memory Database Systems (MMDB), XML Databases, and Data Warehouses. We
presented the challenges involved with each type of system, and how each of the four physical design elements were adapted or utilized to overcome these
challenges. Table \ref{tab:summary} summarizes the results of our survey. From our literature review, we found that the B+tree is the most utilized index
structure, and has been adapted and customized to fit the different needs of other systems. For example, it was made more cache conscious to fit MMDB needs,
and was changed to apply to multidimensions to fit the Data Warehouses needs. Materialized views were found to be heavily utlitized in all the systems we
presented, except in MMDBs. Partitioning was also used in all systems, but was only used for data reloading in MMDBs. Horizontal partitioning is much more
common that vertical partitioning. Clustering is very important in all systems, except for MMDBs. Clustering is especially very important for Data Warehouses
where huge amounts of data are stored.

@@

\begin{table}[!t]
\centering
\scalebox{0.7}{
\begin{tabular}{p{2cm}||p{5cm}|p{4cm}|p{4cm}|p{4cm}}
 & \textbf{Index Structures} & \textbf{Materialized Views} & \textbf{Partitioning} & \textbf{Clustering}\\
\hline\hline
\textbf{RDBMS}& E.g. B+trees, Bitmap Index, Hash Tree & Store results of common complicated queries on disk & Horizontal Partitioning (range, list, hash,
composite) Vertical Partitioning& Group related items together on disk\\
\hline
\textbf{MMDB} & 
-No need to store actual data values in index \linebreak \linebreak 
-Larger node size that is aligned with cache line size \linebreak \linebreak
-E.g. B+tree, CSB+tree, T tree, CST tree & Not highly utilized since processing and main memory access is cheap & Only needed in secondary storage to speed up
reloading in case of a crash\linebreak \linebreak 
-Horizontal \& Vertical partitioning& 
Not needed since random access in main memory costs the same as sequential access\\
\hline
\textbf{XML DB} & Need a number scheme. E.g. B+tree and variations used. & XQuery and XPath results are materialized & Horizontal partitioning based on inlining
or node
type & Based on document order or structural similarity\\
\hline
\textbf{Data Warehouses} & E.g. Universal B+tree, Bitmap index, Projection Index, R*-tree & 
-Reduce query evaluation cost \linebreak
-Reduce view maintenance cost \linebreak 
-Reduce operational costs & 
-Horizontal partitioning based on time dimension \linebreak \linebreak 
-Vertical partitioning of seldom used columns. & Data is
clustered by
nature.\\
\end{tabular}
}
\caption{Summary of the different elements of physical design in the studied systems}
\label{tab:summary}
\end{table}
